### 2019-09-23 supplementary.tex

%&~/tools/latex/preamble-memoir
\usepackage{fontspec}
\usepackage{unicode-math}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\setmathfont[math-style=ISO,bold-style=ISO]{TeX Gyre Pagella Math}
\setmonofont{Inconsolata}

\setsecnumdepth{subsubsection}
\counterwithout{section}{chapter}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}
\date{\today}
\title{Supplementary Material for "The seasonal cycle of upper-ocean turbulence in the Bay of Bengal."}
\begin{document}

\maketitle
\renewcommand{\thefigure}{S\arabic{figure}}
\newcommand{\niw}{_\text{in}}

We present summary time series for all χpods on all moorings in the following figures.

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-NRL1.png}
\captionof{figure}{Daily averaged time series for χpod moorings. NRL1. (left column, top to bottom): Wind stress \(\tau\) (black) and net surface heat flux \(J_q^0\) (color); \(N^2\) at \(\chi\)pod depths; Net shear and \(\chi\)pod depths in black; \(u\); \(v\); EKE and SSH anomaly from AVISO. (right column, top to bottom) T (horizontal lines show \(\chi\)pod depths); S; Near-inertial kinetic energy \(KE\niw\); \(\partial T/\partial z\); \(K_T\); \(J_q^t\) at \(\chi\)pod depths (same colors as in \(N^2\) panel).}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-NRL3.png}
\captionof{figure}{Same as Figure S1 but for NRL3.}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-NRL4.png}
\captionof{figure}{Same as Figure S1 but for NRL4.}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-NRL5.png}
\captionof{figure}{Same as Figure S1 but for NRL5.}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-RAMA-12N-2014.png}
\captionof{figure}{Same as Figure S1 but for RAMA 12N (2014). Shear is not available at the RAMA moorings.}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-RAMA-12N-2015.png}
\captionof{figure}{Same as Figure S5 but for RAMA 12N (2015).}
\end{center}
\end{landscape}

\begin{landscape}
\begin{center}
\includegraphics[width=\linewidth]{images/summaries/summary-RAMA-15N.png}
\captionof{figure}{Same as Figure S5 but for RAMA 15N.}
\end{center}
\end{landscape}
\end{document}